\chapter*{Preface}
In front of you lies the dissertation of \emph{"Optimization of Mobile Communication Systems Using Deep Learning"}. The content is the result of sweat and tears over 3 years. 3 years filled with reading, development and dissementation tasks. But it can safely be said, that it has been 3 years full of wonder and amaze. Mobile Commuication Systems is an incredible engineering achievement, and being allowed to study this combined with the impressive and very recent AI methods have been truly wonderful. 

This dissertation is submitted in partial fulfillment of the requirments for a degree of Doctor of Philosophy, Ph.D at the Technical University of Denmark, May 2020.


\section*{Acknowledgements}

I would like to thank my supervisor Associate Professor, Henrik L. Christiansen for his support, guidance and positive mindset to all ideas put forward. Without this approach to research, many of the novel ideas presented in this dissertation would not have been explored. The past three years could not have been spent in better company.

I would also like to thank my co-supervisor Darko Zibar which introduced me to the environment of academic research. Your early supervison is what made this dissertation a possibility.

To all my other collegeaus at DTU Fotonik over the recent years, I cannot thank you enough. The commeradery has been special and a constant motivation to evolve, both professionally, but also personally.
I would furthermore like to thank the many people who helped proof-read this lengthy document.

Now, enjoy the read and the nice figures

\vspace{4em}


\SignatureAndDate{Jakob Thrane}

